---
title: "Abstract Submission Notes"
author: "by Tim Bender"
subtitle: "useR Conference 2025: Durham, NC"
output: 
  html_document:
    code_download: TRUE
    code_folding: "hide"
    toc: TRUE
    toc_float: TRUE
    toc_depth: "2"
params: 
date: "Updated: 2/19/25"
---
```{r include=FALSE}
library(renv)
library(lubridate)
library(knitr)
library(dplyr)
library(ggplot2)
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

## Dates

```{r include = T, results='hold'}
knitr::opts_chunk$set(message = FALSE)
# build dataset
useR_dates <- data.frame(Date  = strftime(x = as_date(c(rep(ymd(20250303),2),
                                           ymd(20250501),
                                           ymd(20250715),
                                           ymd(20250808)
                                           #ymd(20250810))
                                           )), 
                                  format = "%B %e, %Y", tz = Sys.timezone()), 
                         Event = c("Abstract Submission", 
                                  "Early Bird Registration",
                                  "Program Announcement",
                                  "Regular Registration",
                                  "Event Begins"
                                  #"Event Ends"
                                  )) 

ggplot(data = useR_dates, 
            aes(x = mdy(Date), y = "Date")) + 
  geom_line() +
  geom_label(aes(label = Event), 
             angle = 45, size = 4) +
  geom_label(data = NULL,
             aes(x = Sys.Date(), 
             y = "Date"), 
             label = "Today",
             angle = 45,
             size = 4) 



```



```{r include=TRUE}
knitr::opts_chunk$set(message = FALSE)
```


## Submission Format

-   The title of the tutorial
-   Abstract
-   A brief outline
-   The broad topic it covers
-   The learning goals
-   Time zone preference and time slot (please mention at least three different time zones you are comfortable with) - this is only necessary for the online component of the conference.
-   Length of the tutorial (in minutes or hours)
-   Language in which the tutorial can be taught.
-   The intended audience (beginners, advances, etc.)
-   Pedagogical method of teaching
-   The maximum number of attendees
-   Motivation to teach this tutorial at useR!
    -   Links to the Tutorial materials and/or web page (If there exists).
-   Prerequisites / requirements
-   Material sharing (license), recording consent
-   A brief biography of the instructors

## Sources / Reference

-   [LINK: All about the Abstract Submission](https://rconf.gitlab.io/userknowledgebase/main/tutorials-how-to.html)
-   [LINK: useR 2025 main webpage](https://user2025.r-project.org/)
