---
title: "abstract_brainstorming"
author: "Tim Bender"
date: "`r Sys.Date()`"
output: 
  html_document:
    code_download: TRUE
    code_folding: "hide"
    toc: TRUE
    toc_float: FALSE
    toc_depth: "3"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE)
```

## Presentation Concept

## Similar projects
<!-- - Eclipse 2017 (via [github](https://github.com/RiRam/Eclipse2017))  -->
- The Social Effects of an Awesome Solar Eclipse ([journal article](https://journals.sagepub.com/doi/full/10.1177/09567976221085501), [data hub](https://osf.io/qeku3/))
- [Celestial Programming](https://www.celestialprogramming.com)

## Possible Co-Presenters

#### *The Social Effects of an Awesome Solar Eclipse*
- Sean P. Goldy, University of California, Irvine, Department of Psychological Science
- Paul K. Piff, University of California, Irvine, Department of Psychological Science

#### *Celestial Programming Website*
- Greg Miller: (via [celestialprogramming.com](https://www.celestialprogramming.com), [astronomy stackexchange](https://astronomy.stackexchange.com/users/25729/greg-miller), [asx post](https://astronomy.stackexchange.com/questions/59101/calculation-of-solar-eclipse-progress-based-on-time-and-location#comment137137_59101))

#### *NASA Solar System Ambassadors*
<!-- #### Sinead Borgersen -->
<!-- -   title: Solar System Ambassador -->
<!-- -   agency: NASA -->
<!-- -   contact: [NASA web portal](https://solarsystem.nasa.gov/ambassadors/1574/contact/) -->
<!-- -   website: [NASA Ambassador](https://solarsystem.nasa.gov/ambassadors/1574/) -->
-   Sinead Borgersen: Apex, NC
<!-- #### *Nick Eakes* -->
<!-- -   title: Solar System Ambassador -->
<!-- -   agency: NASA -->
<!-- -   contact: [NASA web portal](https://solarsystem.nasa.gov/ambassadors/1023/contact/) -->
<!-- -   website: [NASA Ambassador](https://solarsystem.nasa.gov/ambassadors/1023/) -->
-   Nick Eakes: Durham, NC
<!-- #### *Christopher Rowe* -->
<!-- -   title: Solar System Ambassador -->
<!-- -   agency: NASA -->
<!-- -   contact: [NASA web portal](https://solarsystem.nasa.gov/ambassadors/2662/contact/) -->
<!-- -   website: [NASA Ambassador](https://solarsystem.nasa.gov/ambassadors/2662/) -->
-   Christopher Rowe: Greensboro, NC

### Map of Speaker Locations

```{r}
library(leaflet)
knitr::opts_chunk$set(warning = FALSE, echo = FALSE)

leaflet(padding = 0, 
        height = "300px", 
        width = "auto") |> 
  addTiles() |> 
  addMarkers(label = c("Sinead Borgersen", "Nick Eakes, Tim Bender", 
                       "Christopher Rowe", "Sean Goldy & Paul Piff"), 
             lng          = c(-78.8508, -78.8986, 
                              -79.7915, -117.8417),
             lat          = c(35.7327,  35.9940, 
                              36.0726, 33.6424)) |>
  expandLimits(lng          = c(-78.8508, -78.8986, 
                              -79.7915, -117.8417)*1.000,
             lat          = c(35.7327,  35.9940, 
                              36.0726, 33.6424)*1.000) |> 
  addTerminator() |> 
  leaflet::addScaleBar(position = "topright") |> 
  leaflet::addSimpleGraticule(zoomIntervals = list(c(0,3,50), 
                                                   c(4,5,5), 
                                                   c(6,20,1)))
```
